---
title: "Slot Machine"
author: "Hands on R programming"
date: "2023-11-07"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The play function will need to do two things. 
First, it will need to randomly generate three symbols; and, 
second, it will need to calculate a prize based on those symbols.

The following function generates three symbols from a group of common slot machine symbols: 
diamonds (DD), sevens (7), triple bars (BBB), double bars (BB), single bars (B), cherries (C), and zeroes (0).
The symbols are selected randomly, and each symbol appears with a different probability:

```{r get_symbols }
get_symbols <- function() {
  wheel <- c("DD", "7", "BBB", "BB", "B", "C", "0")
  sample(wheel, size = 3, replace = TRUE, 
    prob = c(0.03, 0.03, 0.06, 0.1, 0.25, 0.01, 0.52))
}
```

The Manitoba slot machines use the complicated payout scheme shown in Table 9.1. A player will win a prize if he gets:

    Three of the same type of symbol (except for three zeroes)
    Three bars (of mixed variety)
    One or more cherries

Otherwise, the player receives no prize.
the manufacturer claimed they would pay out 92 cents on the dollar. 

A jackpot occurs when a player rolls DD DD DD. Then a player earns $100 doubled three times, which is $800.

Figure 9.2: The score function must distinguish between parallel cases. 
https://rstudio-education.github.io/hopr/images/hopr_0702.png
https://rstudio-education.github.io/hopr/programs.html#fig:subdivide2

https://rstudio-education.github.io/hopr/images/hopr_0703.png
https://rstudio-education.github.io/hopr/programs.html#fig:subdivide3
Figure 9.3: The complete slot machine simulation will involve subtasks that are arranged both in series and in parallel. 

```{r play }
play <- function() {

  # step 1: generate symbols
  symbols <- get_symbols()

  # step 2: display the symbols
  print(symbols)

  # step 3: score the symbols
  score(symbols)
}

```

 we’ve reduced the program to eight simple subtasks:

<1> - Test whether the symbols are three of a kind.
<2> - Test whether the symbols are all bars.
<3> - Look up the prize for three of a kind based on the common symbol.
<4> - Assign a prize of $5.
<5> - Count the number of cherries.
<6> - Count the number of diamonds.
<7> - Calculate a prize based on the number of cherries.
<8> - Adjust the prize for diamonds.

Figure 9.4: score can navigate three cases with two if else decisions. We can also break some of our tasks into two steps. 
https://rstudio-education.github.io/hopr/images/hopr_0704.png
https://rstudio-education.github.io/hopr/programs.html#fig:subdivide4

 One method that I like for checking three of a kind is:
length(unique(symbols) == 1)

```{r score}
score <- function (symbols) {
  # identify case
  same <- symbols[1] == symbols[2] && symbols[2] == symbols[3]
  bars <- symbols %in% c("B", "BB", "BBB")
  
  # get prize
  if (same) {
    payouts <- c("DD" = 100, "7" = 80, "BBB" = 40, "BB" = 25, 
      "B" = 10, "C" = 10, "0" = 0)
    prize <- unname(payouts[symbols[1]])
  } else if (all(bars)) {
    prize <- 5
  } else {
    cherries <- sum(symbols == "C")
    prize <- c(0, 2, 5)[cherries + 1]
  }
  
  # adjust for diamonds
  diamonds <- sum(symbols == "DD")
  prize * 2 ^ diamonds
}
```

```{r efficiency}
#n = 1000
#earn <- c()
#for (i in (1:n)) {
 # earn[i] <- play()
 # }
```


